---
title: "Week 3 exercises"
author: "Brad McNeney"
date: '2019-01-17'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More on matrices *vs* data frames



1. The following simulation function 
simulates $n$ replicates of
an explanatory variable $X$ and a
response variable $Y = \beta X + E$, where $\beta$
is a regression coefficient between $-1$ and $1$
and $E\sim N(0,1)$ is random noise. Run the
code chunk and then use the function to simulate 
one dataset of size $n=1000$ and save the 
result in an object called `dd`.

```{r}
simdat <- function(n) {
  beta <- runif(1,min=-1,max=1)
  x <- rnorm(n)
  y <- beta * x + rnorm(n)
  data.frame(x=x,y=y)
}
######3
dd <- simdat(1000)
```


2. Create a larger dataset by calling `simdat()`
`N=500` times over and stacking the results. 
The larger dataset should have `500*1000` rows and
2 columns.
Call your stacked dataset `bigd1`. 
To create the stacked dataset, 
initialize with `bigd1 <- NULL` and use 
a `for` loop to build up `bigd1` one layer 
at a time. Time this code using the `system.time()` function.
An example use of `system.time()` to time an R 
command, e.g., `x <- rnorm(100000)` is:
```{r}
system.time({
  x <- rnorm(100000) # Could put multiple lines of R code here
})
```
Use the first element of the output (`user` time) as 
your measure of execution time.
```{r}
# bigd1 <- NULL
# for(i in 1:500){
#   bigd1 <- rbind(bigd1,simdat(1000))
# }
#time the code:
system.time({
  bigd1 <- NULL
  for(i in 1:500){
    bigd1 <- rbind(bigd1,simdat(1000))
  } # Could put multiple lines of R code here
})
```
3. Repeat 2, but this time, instead of stacking
the output of `simdat()`, coerce the output
of `simdat()` to a matrix, and stack the 
matrices. Use `system.time()` to time your code
and compare the timing from question (2).
############
Faster than 2
```{r}
system.time({
  bigd1 <- NULL
  for(i in 1:500){
    if(i!=1){
    bigd1 <- rbind(bigd1,as.matrix(simdat(1000)))
    }else{
      bigd1 <- as.matrix(simdat(1000))
    }
  }
})
```
4. Now build `bigd2` by (i) initializing an empty
matrix of appropriate dimension, and (ii) looping 500 times
and inserting simulated datasets of size $n=1000$, coerced to matrices, into successive layers of `bigd2`.
Time this code and compare the timing to that of 
part (3). You may find the following R function
useful:
```{r}
layerInds <- function(layerNum,nrow) {
  ((layerNum-1)*nrow + 1):(layerNum*nrow) 
}
# Example use:
inds <- layerInds(layer=1,nrow=1000)
range(inds)

########################
system.time({
bigd2 <- matrix(data=1:1000000,nrow=500000,ncol=2,
                dimnames = list(c(1:500000),
                                c("x", "y")))

for(i in 1:500){
  bigd2[range(layerInds(i,1000))[1]:range(layerInds(i,1000))[2],]<- as.matrix(simdat(1000))
}
})
```


## Control flow

1. What type of vector does each of the following return?

```{r}
ifelse(TRUE, 1, "no") #double
ifelse(FALSE, 1, "no") #character
ifelse(NA, 1, "no") #logical
```

2. Re-write the following using `switch`

```{r}
IQR_mid <- function(x) mean(quantile(x,c(.25,.75)))
cc <- function(x,method) {
  if(method=="mean") {
    mean(x)
  } else if(method=="median") {
    median(x)
  } else if(method=="IQR_mid") {
    IQR_mid(x)
  } else stop("centring method ",method," not implemented")
}
########
cc <- function(x,method){
  switch(method,mean=mean(x),
         median=median(x),
         IQR_mid=IQR_mid(x),
         stop("centring method ",method," not implemented"))
}
#########
set.seed(123)
x <- c(-3,rnorm(100),1000)
cc(x,"mean")
cc(x,"median")
cc(x,"IQR_mid")
try(cc(x,"cat"))
```

3. Rewrite the following function so that 
it uses a `while()` loop instead of the 
`for()` loop and `break` statement. Your while-approach
will not require the `maxit` upper limit on the 
number of iterations.
```{r}
rtruncNormal <- function(thresh = 2, maxit=1000) {
  x<-NULL
  for(i in 1:maxit) {
    xnew <- rnorm(n=1)
    if(xnew>thresh) {
      break
    }
    x <- c(x,xnew)
  }
  x
}
########
rtruncNormal <- function(thresh = 2){
  x <- NULL
  xnew <- rnorm(n=1)
  while(xnew<thresh){
    x <- c(x,xnew)
    xnew <- rnorm(n=1)
  }
  x
}
set.seed(1234)
rtruncNormal()
```


## Functions

4. The following code chunk is typed into the R Console.
    - What is the output of the function call `f(5)`?
    - What is the enclosing environment of `f()`?
    - What is the enclosing environment of `g()`? #f
    - What search order does R use to find the 
    value of `x` when it is needed in `g()`? inside g->f->global
```{r}
x <- 1
f <- function(y) {
  g <- function(z) {
    (x+z)^2
  }
  g(y)
}
f(5) #36
environment(f)#<environment: R_GlobalEnv>

```


