---
title: "Week 2 Exercises"
author: "Brad McNeney"
output: 
  html_document:
    fig_width: 6
    fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Predict the outcome of the following:
```{r}
c(1, FALSE)
c("a",1)
c(list(1),"a")
c(TRUE,1)
```

2. If `vv <- list(a=1,b=2)`, why doesn't `as.vector(vv)`
work to coerce `vv` into an atomic vector?
->Lists are already vectors 

3. What do `dim()`, `nrow()` and `ncol()` return when applied to a 1-dimensional vector? What about `NROW()` or `NCOL()`?

null, null, null
1, 1

4. What is `dim(cbind(A,A))` if `A = matrix(1:4,2,2)`?
->2 4

5. What do the following return? Understand why. 
```
TRUE | FALSE
c(TRUE,TRUE) & c(FALSE,TRUE)
c(TRUE,TRUE) && c(FALSE,TRUE)
```
TRUE
FALSE TRUE
FALSE


6. What sort of object does `table()` return? What is its type? What attributes does it have? How does the dimensionality change as you tabulate more variables?

table() returns a contingency table of its input variables. It is implemented as an integer vector with class table and dimensions (which makes it act like an array). Its attributes are dim (dimensions) and dimnames (one name for each input column). The dimensions correspond to the number of unique values (factor levels) in each input variable.

7. What happens to a factor when you modify its levels?
How do f2 and f3 differ from f1?

The underlying integer values stay the same, but the levels are changed, making it look like the data has changed.

For f2 and f3 either the order of the factor elements or its levels are being reversed. For f1 both transformations are occurring.

```{r}
f1 <- factor(letters)
levels(f1) <- rev(levels(f1))
f2 <- rev(factor(letters))
f3 <- factor(letters, levels = rev(letters))
```

8. Fix each of the following common data frame subsetting errors:
```{r}
#mtcars[mtcars$cyl = 4, ]
mtcars[mtcars$cyl == 4, ]
#mtcars[-1:4, ]
mtcars[-(1:4) ]
#mtcars[mtcars$cyl <= 5]
mtcars[mtcars$cyl <=5,]
#mtcars[mtcars$cyl == 4 | 6, ]
mtcars[mtcars$cyl %in% c(4,6), ]
```

9. Consider the following data frame and tibble. Suppose
you want to extract the number from the first row and first column and 
multiply this number by the vector `1:3`. How would
you do this for the data frame? How would you do this
for the tibble?
```{r}
dd <- data.frame(x=1:3,y=c("a","b","c"))
tt <- as_tibble(dd)
dd[1][1,1]*c(1:3)
dd$x[1]*c(1:3)
tt$x[1]*c(1:3)

```

10. Read the help file `help("%in%")`. Copy the following data frame into R.
```
students <- data.frame(
           major=c("Stat","Stat","Econ","Bus","Phys","ActSci","ActSci"),
           year=c(3,4,4,5,3,3,4))
```
Use `%in%` to create a logical vector of students who major in Stat
or ActSci and use this to subset `students` to Stat or ActSci majors.

students[students$major %in% c("Stat", "ActSci"),]