---
title: "Stat 360 Week 1 exercises"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. On which line does a get copied in the following example?
-> On line 3
```{r}
library(lobstr)
a <- c(1, 5, 3, 2)
b <- a
# c(obj_addr(a),obj_addr(b))
tracemem(a)
b[[1]] <- 10
# c(obj_addr(a),obj_addr(b))
```

2. In the following code chunk, does `x` get copied? Does `ll` get copied? Does  `ee` get copied?
-> only ll get copied
```{r}
x <- rnorm(100); y <- rnorm(100)
#c(obj_addr(x),obj_addr(y))
cat(tracemem(x), "\n")
cat(tracemem(y), "\n")
ll <- list(x=x,y=y)
cat(tracemem(ll), "\n")
#c(obj_addr(ll))
ll$x <- 1:100
ee <- rlang::env(x=x,y=y)
#cat(tracemem(ee), "\n")
#c(obj_addr(ee))
ee$x <- 1:100
#c(obj_addr(x),obj_addr(y),obj_addr(ll),obj_addr(ee))
```

3. Find the size of the objects x and y in the following code chunk.
Which is smaller? If instead of vectors from 1 to 10 they were vectors
from 1 to 1 million, which would be smaller?
->x is smaller
-> if to 1 million, y is smaller
```{r}
#x <- c(1,2,3,4,5,6,7,8,9,10)
x <- c(1:10)
print(x)
obj_size(x)
y <- 1:10
obj_size(y)
x1 <- c(1:1000000)
obj_size(x1)
y1 <- 1:1000000
obj_size(y1)
```
```{r}
# x <- rnorm(1e6)
# tracemem(x)
# a <- x
# a <- 2
# obj_size(x)
dd <- data.frame(x=rnorm(1e6),y=rnorm(1e6))
system.time({ dd2 <- cbind(dd,dd) })
ref(dd, character = TRUE)
    
mm <- matrix(c(rnorm(1e6),rnorm(1e6)),ncol=2)
system.time({ mm2 <- cbind(mm,mm) })
ref(mm, character = TRUE)
  
```